Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-05-29 manuscript_results.Rmd
update

Example of influenza hospitalizations forecasts and 95% confidence interval.

# Create a data frame with forecasts for plotting
forecast_df <- data.frame(
  Week = factor(1:4, labels = paste0("Week ", 1:4)),  # Label weeks
  Value = c(AUTO_ARIMA_WEEK1_list[[1]]$forecasts[5], 
            AUTO_ARIMA_WEEK2_list[[1]]$forecasts[5], 
            AUTO_ARIMA_WEEK3_list[[1]]$forecasts[5], 
            AUTO_ARIMA_WEEK4_list[[1]]$forecasts[5]),
  Lower = c(AUTO_ARIMA_WEEK1_list[[1]][["0.025"]][5], 
            AUTO_ARIMA_WEEK2_list[[1]][["0.025"]][5], 
            AUTO_ARIMA_WEEK3_list[[1]][["0.025"]][5], 
            AUTO_ARIMA_WEEK4_list[[1]][["0.025"]][5]),
  Upper = c(AUTO_ARIMA_WEEK1_list[[1]][["0.975"]][5], 
            AUTO_ARIMA_WEEK2_list[[1]][["0.975"]][5], 
            AUTO_ARIMA_WEEK3_list[[1]][["0.975"]][5], 
            AUTO_ARIMA_WEEK4_list[[1]][["0.975"]][5])
)

# Plotting forecasts and ribbon
plt00 <- ggplot(forecast_df, aes(x = Week, y = Value, group = 1)) +
  geom_line(color = "black", size = 1) +          # Line connecting forecast values
  geom_point(color = "black", size = 3) +         # Points for each forecast value
  geom_ribbon(aes(ymin = Lower, ymax = Upper),   # Ribbon for confidence intervals
              fill = "gray", alpha = 0.25) +
  labs(title = "Figure 1: Example of probabilistic forecasts",
       subtitle = "The gray ribbon indicates the 95% confidence interval",
       x = "Target Weeks",
       y = "Forecast Values") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 14))

plt00

ggsave(plt00, height = 4, width =7, filename="Fig1.jpg") 
```

@@

Calculate mean weighted interval score for all forecast weeks on each model.

>>>>

@@

Fitting data.

```{r}
i  <- 57
ii <- 8 # state number

my_data = read.csv("NHSH_hospitalizations_seed/treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)

# Convert to Date if not already
list_of_states[[ii]]$target_end_date <- as.Date(list_of_states[[ii]]$target_end_date)

# Define the date range
end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i]-7)
start_date <- end_date - 365*2  # 2 years before

# Subset the data
subset_data <- subset(list_of_states[[ii]], 
                      target_end_date >= start_date & target_end_date <= end_date)

# Plot
plot(subset_data$target_end_date, subset_data$cases, type = "l",
     xlab = "Date", ylab = "Cases", main = "Cases - Last 2 Years")
```

```{r}
# 2) build your forecast_df for i=55, ii=8

i  <- 57
ii <- 8 # state number

forecast_df <- data.frame(
  Dates = c(
    AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
    AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
    AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
    AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
  ),
  Hosp  = c(
    AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
    AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
    AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
    AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
  ),
  Value = c(
    AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
    AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
    AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
    AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
  ),
  Lower = c(
    AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
    AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
    AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
    AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
  ),
  Upper = c(
    AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
    AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
    AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
    AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
  )
)

conceptual_figure = ggplot() +
  # 3a) two-year observed history as a black line (mapped to color for legend)
  geom_point(data = subset_data, 
            aes(x = target_end_date, y = cases, color = "Fitting data"), 
            size = 2.5) +

  # 3b) forecast ribbon (not in legend)
  geom_ribbon(data = forecast_df, 
              aes(x = Dates, ymin = Lower, ymax = Upper), 
              fill = "gray", alpha = 0.4) +

  # 3c) forecast mean (dashed blue line + hollow points)
  geom_point(data = forecast_df, 
            aes(x = Dates, y = Value, color = "Mean forecast"), 
             size = 3) +

  geom_point(data = forecast_df, 
             aes(x = Dates, y = Value, color = "Mean forecast"), 
             shape = 21, fill = "white", size = 2) +

  # 3d) actual post-cutoff cases (mapped to color for legend)
  geom_point(data = forecast_df, 
             aes(x = Dates, y = Hosp, color = "Hospitalizations"), 
             shape = 16, size = 2.5) +

  # Manually define legend entries
  scale_color_manual(name = "Legend",
                     values = c("Fitting data" = "red", 
                                "Mean forecast" = "cyan3",
                                "Hospitalizations" = "black")) +
    scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +

  labs(
    title    = "Example of probabilistic forecasts",
    subtitle = "Gray ribbon represents 95% confidence interval",
    x        = "",
    y        = "Hospitalizations"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1))

conceptual_figure
# Save the combined plot
ggsave("Fig1.jpg", conceptual_figure, width = 7, height = 4, dpi = 600)

```
Calculate mean weighted interval score for all forecast weeks on each model.
